Simulated Structure Factors

### 2018-07-14 Ben_Manuscripts/fourier_transforms/FT_simple_systems.tex
Many new figures and additions to text

@@

  We randomly aligned columns along the z-axis by adding a random displacement
  to each column of points. This simulates a system in which columns are
  uncorrelated.  We shifted coordinates where necessary so that all atoms stayed
  within the unit cell. We held the spacing between each point within each column
  at 3.7 \AA. We created trajectories of 1000 independent
  configurations in order to calculate the average intensity of R-$\pi$. 1000
  independent configurations gives a reasonably converged average, enough to
  observe trends.

  The intensity of R-$\pi$ is equal to the number of scatterers per
  column when we randomly displace columns with respect to each other (See
  Table~\ref{table:randomly_displaced_columns}).  It is independent of the
  number of pores (and hence number of columns) in the system. The error 
  in the calculated intensities are comparable to the magnitude of the 
  intensity because the intensity of each configuration in the trajectory
  fluctuates so much.

@@

@@

@@

@@

  When we add noise in the z direction, the intensity of R-$\pi$ decays while the
  FWHM remains infinite (Figure~\ref{fig:z_noise_columns}).

@@

  \subsection{Finite z-correlations}

@@

>>>>

  \textbf{NEW} There are two ways to set up the pores for this system. The
  first is to make all pores exactly the same; just copies that are translated to
  different pore locations (Figure~\ref{fig:same_rotation}). It is not clear
  whether a perfect crystal of this material would be arranged in this way. The
  second way is to randomly rotate each pore
  (Figure~\ref{fig:different_rotation}). In this work we use the latter.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{same_rotation.png}
  \caption{}\label{fig:same_rotation}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{different_rotation.png}
  \caption{}\label{fig:different_rotation}
  \end{subfigure}
  \caption{A top-down (xy) view of the pore columns. (a) The right pore is an
  exact duplicate of the left including rotation about the z-axis. (b) The right
  pore is a duplicate of the left pore but rotated a random amount about the 
  z-axis.}\label{fig:initial_rotations}
  \end{figure}

@@

  We randomly aligned columns along the z-axis by adding a random displacement
  to each column of points. This simulates a system in which columns are
  uncorrelated.  We wrapped coordinates where necessary so that all atoms stayed
  within the unit cell. We held the spacing between each point within each column
  at 3.7 \AA. We created trajectories of 1000 independent
  configurations in order to calculate the average intensity of R-$\pi$. 1000
  independent configurations gives a reasonably converged average, enough to
  observe trends.

  \textbf{NEW} We varied the degree to which we shifted each column with
  respect to other columns. The most we allowed the columns to shift was one full
  layer. To choose the position of the column, we chose a random displacement
  between 0 and d, where d is the maximum allowed displacement. We varied d
  between 0 and 3.7 \AA.

  When d = 3.7 \AA, the intensity of R-$\pi$ is equal to the number of
  scatterers per column (See Table~\ref{table:randomly_displaced_columns}).  It
  is independent of the number of pores (and hence number of columns) in the
  system. The error in the calculated intensities are comparable to the magnitude
  of the intensity because the intensity of each configuration in the trajectory
  fluctuates so much.

@@

  \textbf{NEW} As we allow the columns to displace less (lower values of d),
  the intensity of R-$\pi$ increases (See Figure~\ref{fig:column_displacement}).
  Once there is no displacement, the intensity of R-$\pi$ matches that of a
  perfect crystal. The intensity of R-$\pi$ in our system is likely boosted by
  correlation between columns.

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{column_displacement.png}
  \caption{As we increase the amount that columns are allowed to displaced relative
  to each other, the intensity of R-$\pi$ decreases.}\label{fig:column_displacement}
  \end{figure}

@@

  \textbf{NEW} There is no major difference in the intensity of R-$\pi$ between
  a perfect crystal and a system with randomly rotated layers. In the z
  direction, R-$\pi$ remains constant since the z components of the points are
  spaced exactly 3.7 \AA apart. The maximum intensity of R-$\pi$ is at ($q_x,
  q_y, q_z$) = (0, 0, 1.7) so there is no contribution from the x or y
  components. Figure~\ref{fig:perfect_v_random_layers} shows a comparison of the
  cross-section of R-$\pi$ along $q_y=0$. On the surface it appears that there is
  a difference between them at high values of $q_y$. However, this is dependent
  on the initial configuration of the perfect crystal system. The system with
  randomly rotated layers will always look as it does if a sufficient number of
  configurations are included in the calculation of the structure factor. The
  perfect system is in agreement with the randomly rotated layer system at low
  values of $q_y$, corresponding to scatteres spaced far apart, however it begins
  to deviate at large $q_y$, where it describes the spatial relationship between
  scatterers in a single pore. In the perfect system, pores are randomly rotated
  with respect to each other (See Figure~\ref{fig:same_rotation}), so the spatial
  relationship changes each time a system is set up. It will not average out since
  each frame is exactly the same, but if enough pores are added, the 2 
  cross-sections will match (See Figure~\ref{fig:perfect_v_random_100pores}). 
  If we create a perfect crystal system where pores are duplicated and translated, but
  not rotated (See Figure~\ref{fig:same_rotation}), the subharmonics
  at large $q_y$ are quite pronounced (Figure~\ref{fig:perfect_crystal_same_rotation})
  due to the increase in ordering in the $q_y$ direction. The peaks are likely
  geometrically related to the spacing between points within pores, but it is 
  not important to get to the bottom of that here.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{perfect_v_random_layers.png} 
  \caption{}\label{fig:perfect_v_random_layers}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{perfect_v_random_100pores.png} 
  \caption{}\label{fig:perfect_v_random_100pores}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{perfect_crystal_same_rotation_qy.png} 
  \caption{}\label{fig:perfect_crystal_same_rotation}
  \end{subfigure}
  \caption{\textbf{NEW} Shown are cross-sections of the 3D structure factor
  along $q_y$ at the $q_z$ value of R-$\pi$. We compare the cross-section for 
  a perfect crystal system and a system where layers are randomly rotated with
  respect to vertically adjacent layers. (a) When systems are made with 4 pores,
  the cross-sections agree at low $q_y$ values, however they diverge at higher
  values of $q_y$. (b) When we build the systems with 100 pores, the cross-sections
  are in close agreement. (c) If we do not rotate the pores in the perfect crystal
  system, there are large spikes at high values of $q_y$ due to order within pores.}\label{fig:perfect_v_layers}
  \end{figure}

@@

  \subsubsection{\textbf{NEW} Perfect Crystal}

  When we add z-direction gaussian noise to the lattice sites in the perfect
  crystal, the intensity of R-$\pi$ drops precipitously as noise is increased
  (Figure~\ref{fig:z_noise_perfect}).  However, the FWHM in the $q_y$ direction
  is not affected by this noise.

  When we add gaussian noise in the x and y directions, the intensity of
  R-$\pi$ does not change, however the FWHM sharpens with increasing noise
  (Figure~\ref{fig:xy_noise_perfect}).

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{z_noise_perfect.png}
  \caption{}\label{fig:z_noise_perfect}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{xy_noise_perfect.png}
  \caption{}\label{fig:xy_noise_perfect}
  \end{subfigure}
  \caption{(a) The intensity of R-$\pi$ decreases as the amount of z-direction
  noise increases. (b) The FWHM decreases as noise in the x and y directions
  increases.}\label{fig:perfect_noise}
  \end{figure}

@@

  When we add noise in the z direction, the intensity of R-$\pi$ decays
  (Figure~\ref{fig:z_noise_columns}).

@@

  \subsubsection{Finite z-correlations}

@@

  \subsubsection{\textbf{NEW} Systems with the "right" amount of noise}

  For a better understanding of the reasons why R-$\pi$ appears as it does in
  our simulated systems, we added a similar amount of noise to the model system
  as that seen in our simulations. Table ~\ref{table:simulation_noise} shows the
  amount of thermal noise in each dimension of the simulated system. We obtained
  those values by calculating the standard deviation in each dimension of the
  center of mass head group location from reference locations. We took the
  reference locations as the average center of mass of each head group. The
  standard deviations reported are only for frames over which we calculated the
  simulated diffraction patterns.

  \begin{table}
  \centering
  \begin{tabular}{c c c c}
  \toprule
  System                         &   $\sigma_x$ ($\AA$) &   $\sigma_y$ ($\AA$) & $\sigma_z$ ($\AA$) \\
  \midrule
  Sandwiched, Ordered            &      0.31            &     0.33             &     0.34           \\
  Parallel Displaced, Ordered    &      0.51            &     0.52             &     0.30           \\
  Sandwiched, Disordered         &      0.41            &     0.57             &     0.32           \\
  Parallel Displaced, Disordered &      0.39            &     0.31             &     0.33           \\
  \bottomrule
  \end{tabular}
  \caption{Thermal noise in each dimension of the center of mass of monomer head groups for 
  simulated LLC systems}\label{table:simulation_noise}
  \end{table}

  Although the head groups do not move much in the equilibrated portion of the
  simulation, they are not arranged perfectly into columns. There is quenched
  disorder present from early in the simulation. To account for this noise, we
  measured the deviation of the center of masses from idealized positions. To
  calculate deviations in the x and y dimensions, we calculated the distance
  between each center of mass and the pore center. The distribution of the radial
  distance between each center of mass and the pore center, for the ordered
  sandwiched configuration, is shown in Figure~\ref{fig:xy_com_distribution}. To
  calculate deviations in the z dimension, we estimated the distance between
  layers to be the average z-box dimension divided by the number of scatterers
  per column (20 in the simulated system). Then we measured the deviation between
  the z position of each center of mass and its corresponding average z-layer
  position. Histograms of the deviations, for the ordered sandwiched
  configuration, are in Figure~\ref{fig:z_com_distribution}. We use the standard
  deviation of each distribution in Figure~\ref{fig:com_distribution} in order to
  place points in our model systems. The standard deviations for all systems
  are given in Table~\ref{table:com_sigmas}.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{xy_com_distribution.png}
  \caption{}\label{fig:xy_com_distribution}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{z_com_distribution.png}
  \caption{}\label{fig:z_com_distribution}
  \end{subfigure}
  \caption{}\label{fig:com_distribution}
  \end{figure}

  \begin{table}[!htb]
  \centering
  \begin{tabular}{c c c}
  \toprule
  System                         &   $\sigma_r$ ($\AA$) &   $\sigma_z$ ($\AA$) \\
  \midrule
  Sandwiched, Ordered            &      2.30            &     1.48             \\
  Parallel Displaced, Ordered    &      2.28            &     1.45             \\
  Sandwiched, Disordered         &      2.93            &     1.58             \\
  Parallel Displaced, Disordered &      2.63            &     1.65             \\
  \bottomrule
  \end{tabular}
  \caption{Thermal noise in each dimension of the center of mass of monomer
  head groups for simulated LLC systems}\label{table:com_sigmas}
  \end{table}

  For a good structural comparison, we set up a model system that mimics the
  ordered sandwiched configuration. We stacked scatterers 4.4 \AA apart and
  created pores with a radius of 6.6 \AA in order to be consistent with the
  values used to calculate the deviations in Table~\ref{table:com_sigmas}. We
  created an initial configuration with noise in the placement of points based on
  the values in Table~\ref{table:com_sigmas}. We made the scatterers correlated
  in the z-direction with a correlation length equal to 20 \AA which is about
  what we've measured in the simulations. Using the initial configuration, we
  created 1000 frames to which we added thermal noise in accordance with that
  measured and recorded in Table~\ref{table:simulation_noise}. The result is a
  center-of-mass trajectory that is configured and behaves similarly to the
  simulation trajectory. We can use that model to tune the various influences
  on R-$\pi$.

  The intensity of R-$\pi$ is strongly dependent on the initial configuration.
  Each time a model system is generated, the initial configuration is different
  than the previous model system. The thermal noise added does not compensate
  for these differences. We ran 40 different model systems with the only 
  difference being the initial randomization. The average R-$\pi$ intensity of the 
  40 trials was 25.7 $\pm$ 9.4. This spread has important implications for the
  reproducibility of our data since there is velocity randomization during the
  initial equilibration. Also, this initial configuration dependence is not 
  very useful for analyzing trends.

  We observe the same R-$\pi$ intensity when each frame of the trajectory is
  randomized. We created a trajectory of 1000 frames. We set up each frame as
  described in the paragraph before last, however with each frame independent. The
  average R-$\pi$ intensity over 1000 frames was 25.4, in close agreement with
  that calculated for the 40 trials of 100 dependent configurations in the
  previous paragraph. We can use this type of setup to analyze trends.

  Allowing columns to displace further with respect to each other causes the
  intensity of R-$\pi$ to decrease (Figure~\ref{fig:ld}). In all of the above
  systems, we placed columns at exactly the same z displacement. Here we allowed
  them to displace from their initial centers as a function of the fraction of
  the distance between layers.

  Qualitatively, we can make R-$\pi$ look the most like experiment when columns
  move independently (See Figure~\ref{fig:rzplots}). As we decrease the
  dependence between columns, the intensity of R-$\pi$ decreases and the
  reflection spreads out in the $q_r$ direction. With any amount of dependence
  between columns, R-$\pi$ sharpens at the center of the reflection.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{rzplot_ld0.png}
  \caption{}\label{fig:rzplot_ld0}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{rzplot_ld33.png}
  \caption{}\label{fig:rzplot_ld33}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{rzplot_ld66.png}
  \caption{}\label{fig:rzplot_ld66}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{rzplot_ld1.png}
  \caption{}\label{fig:rzplot_ld1}
  \end{subfigure}
  \caption{As we allow layers to move more independently, the intensity of
  R-$\pi$ decreases and spreads out in the $q_r$ direction.}\label{fig:rzplots}
  \end{figure}

  Again, R-$\pi$ decreases in response to thermal noise (See 
  Figure~\ref{fig:td}). We varied the amount of thermal noise from a 100 \%
  decrease to a 100 \% increase in noise in each dimension.

  An increased correlation length causes a modest, linear increase in the
  intensity of R-$\pi$ (Figure~\ref{fig:cl}). The effect is not nearly as
  pronounced as z-directional noise and column displacement.

  As we increase the amount of noise in the initial, quenched disordered
  configuration, the intensity of R-$\pi$ decreases (Figure~\ref{fig:IC}).  We
  changed the amount of order in the initial configuration by scaling the noise
  used during placement from -100 \% to 100 \%. We scaled noise in all three
  dimensions, however we know that the maximum intensity of R-$\pi$ will only be
  influenced by noise in the z direction.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{ld.png}
  \caption{}\label{fig:ld}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{td.png}
  \caption{}\label{fig:td}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{cl.png}
  \caption{}\label{fig:cl}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth} 
  \includegraphics[width=\textwidth]{IC.png}
  \caption{}\label{fig:IC}
  \end{subfigure}
  \caption{(a) The intensity of R-$\pi$ decreases as the layers become more
  independent. (b) The intensity of R-$\pi$ decreases when thermal disorder
  decreases. (c) The intensity of R-$\pi$ increases with the correlation length.
  (d) The intensity of R-$\pi$ decreases as more noise is added to the initial
  configuration.}\label{fig:sim_comp}
  \end{figure}
